Seeing Double: Indications From {\it TESS} and {\it Gaia} That PTFO$\,$8-8695b is Not a Planet

### 2020-04-16 paper/ms.tex
update title

\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}

@@

\begin{abstract}
  PTFO$\,$8-8695b is a candidate hot Jupiter in the 7--10$,$Myr old
  Orion-OB1a cluster. It might be the youngest hot Jupiter known.
  We inspected data from TESS and Gaia to clarify whether it is
  actually a planet.
  The TESS lightcurve shows that the dominant variability in this
  system is a sinusoidal modulation with a ``long'' period $P_{\rm
  \ell}=11.96\,$hr, likely caused by stellar rotation.  Also present
  is a complex signal, previously identified as the planet candidate,
  that repeats with a ``short'' period $P_{\rm s}= 10.74\,$hr.
  The ``long'' and ``short'' signals beat every 4.48 days.
  Although there is a dip in the short-period signal, ground-based
  photometry from the past decade shows that the orbital phase of the
  dip seems to have instantaneously jumped, at least once, and perhaps
  twice.
  Planets do not ``jump'' in orbital phase.
  Furthermore, the Gaia data show that PTFO$\,$8-8695 is probably a
  photometric binary, relative to the subgroup of Orion for which
  PTFO$\,$8-8695 is a member.
  Given the available evidence, we believe that PTFO$\,$8-8695 is a
  binary M dwarf system in which one star is showing the ``long''
  rotation signal, and the other star is showing ``transient dipping''
  that has been observed in many young M dwarfs.
  The planetary interpretation seems highly unlikely.
\end{abstract}

@@

% \begin{figure*}[t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=0.9\textwidth]{f6.png}
% 	\end{center}
% 	\vspace{-0.7cm}
% 	\caption{ {\bf Positions, kinematics, and HR diagram of \ptfo.}
%   Members of the 25$\,$Ori-1 subgroup are shown with black circles,
%   and were identified by \citet{kounkel_apogee2_2018} through
%   clustering on Gaia-DR2 and APOGEE data.  The ``neighborhood'' (gray
%   circles) is defined as the group of at most $10^4$ randomly selected
%   non-member stars within 5 standard deviations of the mean right
%   ascension, declination, and parallax.
% 	\label{fig:gaia}
% 	}
% \end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.8\textwidth]{f6.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf HR diagram of \ptfo\ and members of 25$\,$Ori-1.}
  Members of the 25$\,$Ori-1 group (black circles) were identified by
  \citet{kounkel_apogee2_2018} through clustering on six-dimensional
  Gaia-DR2 and APOGEE-2 data.  We defined the reference
  ``neighborhood'' (gray circles) as the group of at most $10^4$
  randomly selected non-member stars within 5 standard deviations of
  the mean 25$\,$Ori-1 right ascension, declination, and parallax.
  \ptfo\ seems likely to be a photometric binary.
  ``$G$'' is the Gaia broadband, $Bp$ is Gaia blue, and $Rp$ is Gaia red.
  $\omega_{\rm as}$ is the parallax in arcseconds.
	\label{fig:gaia}
	}
\end{figure*}

>>>>

%\title{Against the Planetary Interpretation of PTFO$\,$8-8695b}
% \title{Seeing Double: Still Against the Planetary Interpretation of PTFO$\,$8-8695b}
% \title{Seeing Double: TESS and Gaia Show That PTFO$\,$8-8695b is Unlikely a
% Planet}
\title{Seeing Double: Indications From {\it TESS} and {\it Gaia} That
PTFO$\,$8-8695b is Not a Planet}

@@

\begin{abstract}
  PTFO$\,$8-8695b is a candidate hot Jupiter in the 7--10 million year old
  Orion-OB1a cluster. It might be the youngest hot Jupiter known.
  We inspected data from TESS and Gaia to clarify whether it is
  actually a planet.
  The TESS lightcurve shows that the dominant variability in this
  system is a sinusoid with a ``long'' period $P_{\rm
  \ell}=11.96\,$hr, previously thought to be a systematic, but in fact likely caused by stellar rotation.  Also present
  is a complex signal, previously identified as the planet candidate,
  that repeats with a ``short'' period $P_{\rm s}= 10.74\,$hr.
  The two signals beat every 4.48 days.
  Although there is a dip in the short-period signal, ground-based
  photometry from the past decade shows that the orbital phase of the
  dip seems to have instantaneously jumped, at least once, and perhaps
  twice.
  Planets do not ``jump'' in orbital phase.
  Furthermore, the Gaia data show that PTFO$\,$8-8695 is probably a
  photometric binary.
  Given the available evidence, we believe that PTFO$\,$8-8695 is a
  binary M dwarf in which one star shows the ``long''
  rotation signal, and the other star is showing ``transient dipping''
  that has also been observed in other young M dwarfs.
  The planetary interpretation seems quite unlikely.
\end{abstract}

@@

% \begin{figure*}[t]
% 	\begin{center}
% 		\leavevmode
% 		\includegraphics[width=0.9\textwidth]{f6.png}
% 	\end{center}
% 	\vspace{-0.7cm}
% 	\caption{ {\bf Positions, kinematics, and HR diagram of \ptfo.}
%   Members of the 25$\,$Ori-1 subgroup are shown with black circles,
%   and were identified by \citet{kounkel_apogee2_2018} through
%   clustering on Gaia-DR2 and APOGEE data.  The ``neighborhood'' (gray
%   circles) is defined as the group of at most $10^4$ randomly selected
%   non-member stars within 5 standard deviations of the mean right
%   ascension, declination, and parallax.
% 	\label{fig:gaia}
% 	}
% \end{figure*}
\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.8\textwidth]{f6.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{ {\bf HR diagram of \ptfo\ and members of 25$\,$Ori-1.}
  Members of the 25$\,$Ori-1 group (black circles) were identified by
  \citet{kounkel_apogee2_2018} through clustering on six-dimensional
  Gaia-DR2 and APOGEE-2 data.  We defined the reference
  ``neighborhood'' (gray circles) as the group of at most $10^4$
  randomly selected non-member stars within 5 standard deviations of
  the mean 25$\,$Ori-1 right ascension, declination, and parallax.
  \ptfo\ seems likely to be a photometric binary.
  $G$ is the Gaia broadband, $Bp$ is Gaia blue, $Rp$ is Gaia red, and
  $\omega_{\rm as}$ is the parallax in arcseconds.
	\label{fig:gaia}
	}
\end{figure*}
